Introduction

### 2025-12-02 Intro.Rmd
Add files via upload

#bank_qda <- qda(y ~., data = bank_full, subset = train)
```

@@

@@

```{r}
library(gam)
#bank_gam <- gam(y ~ , data = bank_train) # basis functions

## Boosting Decision Trees
require post-hoc explanation

Potential model Mixed effects model (attribute set variables to have a random effect)

>>>>

#bank_qda <- qda(y ~., data = bank_full, subset = train)

set.seed(123)
train_index <- sample(seq_len(nrow(bank_full)), size = 0.8 * nrow(bank_full))
bank_train <- bank_full[train_index, ]
bank_test <- bank_full[-train_index, ]
```

```{r}
lda_model <- lda(y ~ ., data = bank_train)
lda_model
```

```{r}
lda_pred <- predict(lda_model, newdata = bank_test)$class
lda_actual <- bank_test$y
```

```{r}
lda_accuracy <- mean(lda_pred == lda_actual)
lda_error <- 1 - lda_accuracy
cat("LDA Test Accuracy:", round(lda_accuracy * 100, 2), "%\n")
cat("LDA Test Error:", round(lda_error * 100, 2), "%\n")
```

```{r}
qda_model <- qda(y ~ ., data = bank_train)
qda_model
```

```{r}
qda_pred <- predict(qda_model, newdata = bank_test)$class
qda_actual <- bank_test$y
```

```{r}
qda_accuracy <- mean(qda_pred == qda_actual)
qda_error <- 1 - qda_accuracy
cat("QDA Test Accuracy:", round(qda_accuracy * 100, 2), "%\n")
cat("QDA Test Error:", round(qda_error * 100, 2), "%\n")
```

@@

```{r}
numeric_vars <- bank_full %>%
select(-y) %>%
select(where(is.numeric)) %>%
names()
train_x <- bank_train[, numeric_vars]
test_x <- bank_test[, numeric_vars]
```

```{r}
train_x_scaled <- scale(train_x)
test_x_scaled <- scale(
  test_x,
  center = attr(train_x_scaled, "scaled:center"),
  scale  = attr(train_x_scaled, "scaled:scale")
)
train_y <- bank_train$y
test_y  <- bank_test$y
```

```{r}
knn_pred <- knn(train = train_x_scaled,
test = test_x_scaled,
cl = train_y,
k = 5)
```

```{r}
knn_accuracy <- mean(knn_pred == test_y)
knn_error <- 1 - knn_accuracy
cat("KNN Test Accuracy:", round(knn_accuracy * 100, 2), "%\n")
cat("KNN Test Error:", round(knn_error * 100, 2), "%\n")
```

```{r}
k_values <- 1:25
cv_accuracy <- numeric(length(k_values))
folds <- createFolds(train_y, k = 5)
for (i in seq_along(k_values)) {
k <- k_values[i]
fold_acc <- c()
for (f in folds) {
fold_train_x <- train_x_scaled[-f, ]
fold_test_x <- train_x_scaled[f, ]
fold_train_y <- train_y[-f]
fold_test_y <- train_y[f]
pred <- knn(
  train = fold_train_x,
  test  = fold_test_x,
  cl    = fold_train_y,
  k     = k
)

fold_acc <- c(fold_acc, mean(pred == fold_test_y))
}
cv_accuracy[i] <- mean(fold_acc)
}
best_k <- k_values[which.max(cv_accuracy)]
best_k
```

```{r}
plot(k_values, cv_accuracy, type = "b",
xlab = "K Value", ylab = "Cross-Validated Accuracy",
main = "K Optimization for KNN")
abline(v = best_k, col = "red", lty = 2)
```

```{r}
knn_best_pred <- knn(
train = train_x_scaled,
test = test_x_scaled,
cl = train_y,
k = best_k
)
final_accuracy <- mean(knn_best_pred == test_y)
final_error <- 1 - final_accuracy
cat("Best K:", best_k, "\n")
cat("Test Accuracy with Best K:", round(final_accuracy * 100, 2), "%\n")
cat("Test Error with Best K:", round(final_error * 100, 2), "%\n")
```

@@

```{r}
library(gam)
#bank_gam <- gam(y ~ , data = bank_train) # basis functions
```

```{r}
gam_model <- gam(
y ~ s(age) + s(balance) + s(duration) + s(campaign) + s(pdays),
family = binomial,
data = bank_train
)
summary(gam_model)
```

```{r}
gam_prob <- predict(gam_model, newdata = bank_test, type = "response")
gam_pred <- ifelse(gam_prob > 0.5, "yes", "no")
gam_pred <- as.factor(gam_pred)
```

```{r}
actual <- bank_test$y
gam_accuracy <- mean(gam_pred == actual)
gam_error <- 1 - gam_accuracy
cat("GAM Test Accuracy:", round(gam_accuracy * 100, 2), "%\n")
cat("GAM Test Error:", round(gam_error * 100, 2), "%\n")
```

## Boosting Decision Trees
require post-hoc explanation
```{r}
library(gbm)
```
```{r}
bank_train$y_num <- ifelse(bank_train$y == "yes", 1, 0)
bank_test$y_num <- ifelse(bank_test$y == "yes", 1, 0)
set.seed(123)
boost_model <- gbm(
formula = y_num ~ . - y,
data = bank_train,
distribution = "bernoulli",
n.trees = 1000,
interaction.depth = 3,
shrinkage = 0.01,
n.minobsinnode = 10,
verbose = FALSE
)
```

```{r}
boost_prob <- predict(boost_model, newdata = bank_test, n.trees = 1000, type = "response")
boost_pred <- ifelse(boost_prob > 0.5, "yes", "no")
```

```{r}
boost_accuracy <- mean(boost_pred == bank_test$y)
boost_error <- 1 - boost_accuracy
cat("Boosting Test Accuracy:", round(boost_accuracy * 100, 2), "%\n")
cat("Boosting Test Error:", round(boost_error * 100, 2), "%\n")
summary(boost_model)
```
Post-hoc explanation: Boosting combines the small decision trees to not make the same mistake the previous one has made. Which creates a very good and accurate model.

Ω
Potential model Mixed effects model (attribute set variables to have a random effect)
